%=========================================================================
% (c) 2014, 2015 Josef Lusticky

\chapter{Introduction}
In recent years not only global Internet traffic is increasing, but also middle and small enterprises
need bandwidth and latency comparable to what ISPs used a few years ago.
Whith the start of incorporating the cloud philosophy, where every device and data is accessible from everywhere,
the traffic within the enterprises' campus network increased as well.

10Gbit software routing using GNU/Linux provides a financially moderate alternative to hardware routers.




40Gbit and 100Gbit over Ethernet - IEEE 802.3ba ratified in 2010~\cite{ieee-802.3ba}.
10GBASE-T (4-pair copper Cat.8 twisted-pair) ratified in 2013.
8P8C modular connector (RJ45).

Then you have QSFP, which is a quad SFP [connection].
So to get to 40 GbE today, you take four lanes of [10 Gbps] and run them together.
You can get to 100 GbE by taking 10 lanes of 10 Gbps.

---

Protocol offload engines (ISO/OSI layers 2-4 implemented in HW)
on Ethernet - %http://www.ics.uci.edu/~ccgrid11/files/ccgrid11-ib-hse_last.pdf
Interrupt coalescing, Jumbo frames, HW checksum engines, Segmentation Offload engines

Accelerators:
TCP Offload Engine - HW acceleration of the entire TCP/IP stack
Internet Wide-Area RDMA Protocol (iWARP)

Converged Ethernet - lossless ethernet for datacenters


----

TCP Segmentation Offload
Zero-copy principle
RDMA
I/O Acceleration Technology


\input{40gbe.tex}

\input{linux.tex}

\input{analysis.tex}

\input{setup.tex}

\input{measurements.tex}


\chapter{Conclusion}
40Gbit software routing using GNU/Linux provides a reasonable alternative to expensive hardware routers.
The main bottleneck is. %TODO
