%=========================================================================
% (c) 2014, 2015 Josef Lusticky

\section{Server configuration}\label{sec:setup-server}
The general system configuration used by default in all measurements
consists of disabling Netfilter, disabling SELinux, changing the scaling governor, disabling rp\_filter and
configuring the IP addresses according to the scheme described in section~\ref{sec:setup-hardware}.
Additionally, the system should run no unused services - CentOS~7 runs Postfix, Avahi daemon, Polkitd, Tuned, Crond,
NetworkManger, dbus, crond, rsyslogd and auditd by default.
It is recommended to disable all of these services for the performance reasons.
Note that dbus may still be activated if required by any of the user-space application the system runs.
Beware that disabling NetworkManager may cause no network connectivity.
If the system uses DHCP to obtain an IP address, the {\it{ifcfg}} network scripts or
the {\it{/etc/rc.local}} script can be used for this purpose.
Execution of {\it{dhclient}} can be added to the {\it{/etc/rc.local}} file and
the file must be made executable by running {\it{chmod +x /etc/rc.local}}.
The {\it{dhclient}} daemon should be killed after it obtained an IP address
to avoid sending DHCP Request packets on other interfaces.
The services can be disable using {\it{systemctl}} command, as the following listing shows.
\begin{lstlisting}
systemctl disable postfix avahi-daemon tuned crond polkit NetworkManager auditd dbus cron rsyslog
reboot
\end{lstlisting}
The {\it{ps uax}} command can be further used to list running processes on the system.
\\

The measurements were executed under the following conditions:
\\
Netfilter was disabled completely:
\begin{lstlisting}[language=TeX]
systemctl stop firewalld
systemctl disable firewalld  # do not execute firewalld on boot
\end{lstlisting}
SELinux was disabled by changing the SELINUX variable to {\it{disabled}} in /etc/sysconfig/selinux
(default is {\it{Enforcing}}). The system must reboot to apply:
\begin{lstlisting}
vi /etc/sysconfig/selinux
	SELINUX=disabled
reboot
\end{lstlisting}
The scaling governor was changed for each CPU to {\it{performance}} (default is {\it{powersave}}):
\begin{lstlisting}
echo performance | tee /sys/devices/system/cpu/cpu[0-9]*/cpufreq/scaling_governor
\end{lstlisting}
The rp\_filter was disabled on all interfaces (it should be disabled at least on the forwarding interfaces):
\begin{lstlisting}[language=TeX]
echo 0 | tee /proc/sys/net/ipv4/conf/*/rp_filter
\end{lstlisting}
It is suggested to disable the rp\_filter at boot time by putting the following lines to the {\it{/etc/sysctl.conf}} file:
\begin{lstlisting}[language=TeX]
net.ipv4.conf.all.rp_filter=0
net.ipv4.conf.default.rp_filter=0
net.ipv4.conf.lo.rp_filter=0
net.ipv4.conf.enp129s0.rp_filter=0        # forwarding interface 1
net.ipv4.conf.enp129s0d1.rp_filter=0      # forwarding interface 2
\end{lstlisting}
IPv6 was enabled on all interfaces (it must be enabled at least on the forwarding interfaces):
\begin{lstlisting}
echo 0 > /proc/sys/net/ipv6/conf/all/disable_ipv6
\end{lstlisting}
IPv4 forwarding was enabled:
\begin{lstlisting}
echo 1 > /proc/sys/net/ipv4/ip_forward
\end{lstlisting}
IPv6 forwarding was enabled on all interfaces (it must be enabled at least on the forwarding interfaces):
\begin{lstlisting}
echo 1 > /proc/sys/net/ipv6/conf/all/forwarding
\end{lstlisting}
IPv4 neighbours were set:
\begin{lstlisting}
ip neigh add 192.0.2.2 lladdr 00:10:94:00:00:01 dev enp6s0d1
ip neigh add 192.0.2.6 lladdr 00:10:94:00:00:02 dev enp6s0
\end{lstlisting}
\newpage
IPv6 neighbours were set:
\begin{lstlisting}
ip -6 neigh add 2001:db8:1::2 lladdr 00:10:94:00:00:03 dev enp6s0d1
ip -6 neigh add 2001:db8:2::6 lladdr 00:10:94:00:00:04 dev enp6s0
\end{lstlisting}
IPv4 addresses were assigned:
\begin{lstlisting}
ip addr add 192.0.2.1/30 broadcast 192.0.2.3 dev enp6s0d1
ip addr add 192.0.2.5/30 broadcast 192.0.2.7 dev enp6s0
\end{lstlisting}
IPv6 addresses were assigned:
\begin{lstlisting}
ip -6 addr add 2001:db8:1::1/64 dev enp6s0d1
ip -6 addr add 2001:db8:2::5/64 dev enp6s0
\end{lstlisting}

The routing performance of the upstream Linux kernel version 4.0.2 was further measured.
The ELRepo was added to available repositories and the kernel was installed.
The instructions to add the ELRepo repository are provided by the elrepo.org site.\footnote{\url{http://elrepo.org/}}
Afterwards, the {\it{kernel-ml}} package was installed:
\begin{lstlisting}
yum --enablerepo=elrepo-kernel install kernel-ml
\end{lstlisting}
The kernel can be set as default in the bootloader configuration.
The following command prints all available kernels on the system:
\begin{lstlisting}[language=TeX]
grep "submenu\|^\menuentry" /boot/grub2/grub.cfg | cut -d "'" -f2
	CentOS Linux, with Linux 4.0.2-1.el7.elrepo.x86_64
	CentOS Linux, with Linux 3.10.0-123.13.1.el7.x86_64
	CentOS Linux, with Linux 0-rescue-f8351e2baaac42a285a6443a1f777333
\end{lstlisting}
The 4.0.2 kernel can be set as default by changing the configuration of grub2:
\begin{lstlisting}[language=TeX]
grub2-set-default 'CentOS Linux, with Linux 4.0.2-1.el7.elrepo.x86_64'
\end{lstlisting}

The routes announced in public BGP were imported to the kernel's FIB to perform additional measurements.
Appendix~\ref{app:bgp} describes the step-by-step instructions on how to obtain the BGP table and import it to the FIB.
Any additional system settings for a particular measurement are described next to the result.
