%=========================================================================
% (c) 2014, 2015 Josef Lusticky

\section{Software and firmware}
Base CentOS 7 was installed on the server.
The operating system features Linux kernel based on version 3.10 -
the installed version is 3.10.0-123.20.1.el7.x86\_64.
The operating system was updated with all updates avaliable as of 20th March 2015.
The upstream kernel version 3.19.2 was additionally installed from the ELRepo repository~\cite{elrepo-kernel-ml}.

The Linux kernel detects the Mellanox ConnectX-3 EN card automatically and loads the mlx4\_core and mlx4\_en module.
The mlx4\_core module prints the detected PCI-Express link parameters to the kernel's message buffer.
The buffer can be viewed using the dmesg command and its partial output is shown bellow:
\begin{lstlisting}[language=TeX]
mlx4_core 0000:06:00.0: PCIe link speed is 8.0GT/s, device supports 8.0GT/s
mlx4_core 0000:06:00.0: PCIe link width is x8, device supports x8
\end{lstlisting}
The mlx4\_core module further registers interrupts and prints the assigned IRQ numbers for each queue
to the kernel's message buffer:
\begin{lstlisting}[language=TeX]
mlx4_core 0000:06:00.0: irq 61 for MSI/MSI-X
mlx4_core 0000:06:00.0: irq 62 for MSI/MSI-X
...
mlx4_core 0000:06:00.0: irq 90 for MSI/MSI-X
\end{lstlisting}

The driver uses either MSI or MSI-X feature of the PCI-Express bus, as described in section~\ref{sec:40gbe-throughput}.
The MSI-X feature is used automatically if the system supports it, otherwise the adapter uses MSI.
The {\it{lspci -vv}} command can be used to check whether MSI-X is used -
the MSI-X capability is followed by an Enable flag which is followed with either "+" (enabled)
or "-" (disabled).
Listing~\ref{lst:setup-lspci} shows partial output of lspci for the Mellanox ConnectX-3 EN adapter.
The system supports MSI-X and the adapter is configured to use it.
\begin{lstlisting}[language=TeX,label={lst:setup-lspci},caption={Partial output of lspci -vv for Mellanox ConnectX-3 EN}]
06:00.0 Ethernet controller: Mellanox Technologies MT27500 Family [ConnectX-3]
		...
		Capabilities: [9c] MSI-X: Enable+ Count=128 Masked-
				...
				LnkCap: Port #8, Speed 8GT/s, Width x8, ASPM L0s, Exit Latency L0s unlimited, L1 unlimited
				...
\end{lstlisting}

Apart from the NIC driver, which is part of kernel package, the Mellanox ConnectX-3 adpater uses its own firmware.
The firmware was updated to version 2.32.5100, which is the latest version available as of 10th January 2015.
The firmware update procedure is described in appendix~\ref{app:firmware}.

%By default, the driver uses adaptive interrupt moderation for the receive path,
%which adjusts the moderation time to the traffic pattern~\cite{mellanox-user-manual}.
