%=========================================================================
% (c) 2014, 2015 Josef Lusticky

\subsection{Ethtool settings}\label{sub:analysis-settings-ethtool}
Ethtool is a standard Linux utility for manipulating network drivers and hardware, particularly for
wired Ethernet devices.
Supported offload features can be displayed using ethtool -{}-show-offload {\it{ifname}}.
Listing~\ref{lst:analysis-ethtool-offload} shows the output of ethtool -{}-show-offload eth0,
where eth0 is the Mellanox ConnectX-3 adapter.

\begin{lstlisting}[caption={Output of ethtool -{}-show-offload for Mellanox ConnectX3 adapter},label={lst:analysis-ethtool-offload}]
rx-checksumming: on
tx-checksumming: on
	tx-checksum-ipv4: on
	tx-checksum-ip-generic: off [fixed]
	tx-checksum-ipv6: on
	tx-checksum-fcoe-crc: off [fixed]
	tx-checksum-sctp: off [fixed]
scatter-gather: on
	tx-scatter-gather: on
	tx-scatter-gather-fraglist: off [fixed]
tcp-segmentation-offload: on
	tx-tcp-segmentation: on
	tx-tcp-ecn-segmentation: off [fixed]
	tx-tcp6-segmentation: on
udp-fragmentation-offload: off [fixed]
generic-segmentation-offload: on
generic-receive-offload: on
large-receive-offload: off [fixed]
rx-vlan-offload: on [fixed]
tx-vlan-offload: on [fixed]
ntuple-filters: off [fixed]
receive-hashing: on
highdma: on [fixed]
rx-vlan-filter: on [fixed]
vlan-challenged: off [fixed]
tx-lockless: off [fixed]
netns-local: off [fixed]
tx-gso-robust: off [fixed]
tx-fcoe-segmentation: off [fixed]
tx-gre-segmentation: off [fixed]
tx-ipip-segmentation: off [fixed]
tx-sit-segmentation: off [fixed]
tx-udp_tnl-segmentation: off [fixed]
tx-mpls-segmentation: off [fixed]
fcoe-mtu: off [fixed]
tx-nocache-copy: on
loopback: off
rx-fcs: off [fixed]
rx-all: off [fixed]
tx-vlan-stag-hw-insert: off [fixed]
rx-vlan-stag-hw-parse: off [fixed]
rx-vlan-stag-filter: off [fixed]
\end{lstlisting}

Selected offload features can be enabled by ethtool -{}-offload {\it{devname}} {\it{feature}} on,
or disabled by ethtool -{}-offload {\it{devname}} {\it{feature}} off.
The features listed as [fixed] cannot be changed on this particular NIC.
The rest of the features can be changed, but the feature name differs when executing ethtool -{}-offload.
For example, the rx-checksumming feature is turned on by ethtool -{}-offload eth0 {\it{rx}} on.
Similarly, the scatter-gather feature is turned on by ethtool -{}-offload eth0 {\it{sg}} on.
For more information about using ethtool, the man page of ethool(8) should be consulted.

Listing~\ref{lst:analysis-ethtool-offload} shows that all supported offload features not marked as [fixed] are enabled,
except the {\it{loopback}} option.
However, the option is not used when processing network traffic between two hosts~\cite{kernel-source}.
This is the default configuration in the CentOS~7 operating system with the Mellanox ConnectX-3 EN adapter.
